### 2020-08-03 outputs/documents/paper.Rmd
Update first four pages

---
output: 
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: false
    toc: false
    fig_caption: true
    latex_engine: pdflatex
    template: templates/svm-latex-ms.tex
bibliography: "references.bib"
header-includes:
  -  \usepackage{hyperref}
biblio-style: apalike
title: "Consistency scores in text data"
thanks: "Thank you to X, Y and Z for helpful comments. A Shiny application for interactive and small-scale examples is available: https://kelichiu.shinyapps.io/Arianna/. An R package for larger scale applications is available: **HERE**. Our code and datasets are available: https://github.com/RohanAlexander/consistency_scores_in_text_datasets. Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: rohan.alexander@utoronto.ca."
author:
- name: Ke-Li Chiu
  affiliation: University of Toronto
- name: Rohan Alexander
  affiliation: University of Toronto
abstract: "In this paper we introduce a process to clean the text extracted from PDFs using various methods from natural language processing. Our approach compares originally extracted text with the text generated from, or expected by, these methods using earlier text as stimulus. To guide this process, we introduce the notion of a consistency score, which refers to the proportion of text that is unchanged by the model. This is used to monitor changes during the cleaning process, and to compare the messiness of different texts. The methods that we consider are: n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3. We illustrate our process on text from the Canadian Hansard and introduce both a Shiny application and an R package to make our process easier for others to adopt."
keywords: "text-as-data; natural language processing; quantitative analysis."
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 12pt
# spacing: double
endnote: no
graphics: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

When we think of quantitative analysis, we may like to think that our job is to 'let the data speak'. But this is rarely the case in practise. Datasets can have errors, be biased, incomplete, or messy. In any case, it is the underlying statistical process of which the dataset is an artifact that is typically of interest. In order to use statistical models to understand that process, we typically need to clean and prepare the dataset in some way. This is especially the case when we work with text data. But this cleaning and preparation requires us to make many decisions. To what extent should we correct obvious errors? What about slightly-less-obvious errors? Although cleaning and preparation is a necessary step, we may be concerned about the extent to which have we introduced new errors, and the possibility that we have made decisions that have affected, or even driven, our results.

In this paper we introduce the concept of consistency in a text corpus. Consistency refers to the proportion of words that are able to be forecast by a statistical model, based on preceding words and surrounding context. Further, we define internal consistency as when the model is trained on the corpus itself, and external consistency as when the model is trained on a more general corpus. Together, these concepts provide a guide to the cleanliness and consistency of a text dataset. This can be important when deciding whether a dataset is fit for purpose; when carrying out data cleaning and preparation tasks; and as a comparison between datasets.

To provide an example, consider the sentence, 'the cat in the\dots'. A child who has read this book could tell you that the next word should be 'hat'. Hence if the sentence was actually 'the cat in the bat', then that child would know something was likely wrong. The consistency score would likely be lower than if the sentence were 'the cat in the hat'. After we correct this error, the consistency score would likely increase. By examining how the consistency scores evolve in response to changes made to the text during the data preparation and cleaning stages we can better understand the effect of the changes. Including consistency scores and their evolution when text corpuses are shared allows researchers to be more transparent about their corpus. And finally, the use of consistency scores allows for an increased level of automation in the cleaning process.

We consider various natural language processing methods. These include n-grams, Google's Word2vec, Stanford's GloVe, and OpenAI's GPT-2 and GPT-3. The n-gram approach involves identifying two, three, or more, words that are commonly found together. Here think of a two-gram involving the word 'good' 'good morning'. At scale these can identify missing or unusual words, and work quickly, but they lack nuance. For instance an equally reasonable two-gram involving the word 'good' is 'good work'. For that reason we consider pre-trained word embedding models. These include Google's Word2vec and Stanford's GloVe, which place each word in a multi-dimensional space such that distance between words can illustrate their relationship. We also consider pre-trained generative models such as OpenAI's GPT-2 and GPT-3. GPT-2 and GPT-3 are pre-trained generative unsupervised language models. They differ in the number of parameters. Generative Pretrained Transformer (GPT)-2 has 1.5 billion parameters and GPT-3 has more than 175 billion parameters. While GPT-2 is more limited, it can be run on smaller machines, whereas GPT-3 cannot and is accessed via an API.

We apply our approach to the Canadian Hansard. The Canadian Hansard was digitised by @beelen2017digitization. This was a process by which PDFs were scanned, put through optical character recognition (OCR), and then corrected to a reasonable extent. The dataset is extensive, fit-for-purpose, and, appropriately, it has been used considerably, for instance @rheault2020word. However, there are many issues with the dataset in terms of reading it. We focus on one particular year - **(Rohan pick one)** - and show that our approach can be used to relatively quickly improve the quality of the dataset in a reproducible and consistent manner.

The remainder of our paper is structured as follows: **(Rohan add structure)**. Additionally, we construct a Shiny app available at: **(Rohan add link)**. That app computes internal and external consistency scores for corpus excerpts, and have developed an R Package that allows our approach to be used on larger datasets, which is available at: **(Rohan add link)**.

# Background

## Language Model

Given the nature of human languages, some combinations of words tend to occur more frequently than others. Think of 'good', which is more often followed by 'morning', than 'duck'. As such, we could consider English text production as a conditional probability, $\mbox{Pr}(w_{k} | w^{k-1}_{1})$, where $k$ is the number of words in a sentence, $w_{k}$ is the predicted word, and $w^{k-1}_{1}$ is the history of the word occurring in a sequence [@brown1992class]. In this way, the generation of some prediction, $w_{k}$, is based on the history, $w^{k-1}_{1}$. This is the underlying principle of all language models. Essentially, a language model is a probability distribution over sequences of words **(Ke-Li add ref)**. The goal of statistical language modeling is to estimate probability distributions over different linguistic units — words, sentences, and even documents **(Ke-Li add ref)**. However, this is difficult as language is categorical. If we consider each word in a vocabulary as a category, then the dimensionality of language becomes large [@rosenfeld2000two]. The reason that there is scuh a variety of statistical language models is that there are various ways of dealing with this fundamental problem.

## PDF text extraction and n-grams

The Portable Document Format (PDF) is typically used to render documents in a relatively  fixed way. PDFs contain a lot of information, but often that information is not able to be analysed directly by statistical language models. The information first needs to be extracted from the PDF, often by optical character recognition (OCR), and then cleaned, prepared, and made into some type of dataset. One issue in this process is that OCR is not perfect. Hence there is a need to correct the output before it is analysed. One way to do this is to correct errors one-by-one and independent of context. We all know that 'tbe' is wrong, and should be 'the'. But statistical language models have the advantage of correction that can occur within context. This allows significant improvements, for instance sometimes 'bat' should be 'hat' and sometimes it really should be 'bat' **(Ke-Li add ref)**.

The foundation of an n-gram language model is the conditional probability set-up introduced above. An n-gram model is a probabilistic language model that predicts the next word in a sequence of words **(Ke-Li add ref)**. The $n$ in n-gram referes to the number of words in that sequence. Consider the following excerpt from *Jane Eyre*:

> 'We had been wandering'.

'We' is a uni-gram, 'We had' is a bi-gram, 'We had been' is a tri-gram, and 'We had been wandering' is a 4-gram. Notice that the two tri-grams in this excerpt, 'We had been', and 'had been wandering', overlap. The use of n-gram models enables us to assign probabilities to both the next sequence of words and just the next word. For instance, consider the two sentences: 'We had been wandering', and 'We had been wangling'. The former is likely to be more frequently encountered in a training corpus. Thus, an n-gram would assign a higher probability to the next word being 'wandering' than 'wangling', given the sequence 'We had been'.

To predict the next word, we have to take the sequence of preceding words into account, which requires knowing the probability of the sequence of words. The probability of a sequence appearing in a corpus follows the chain rule:

$$\mbox{Pr(We,had,been,wandering)} = \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|We,had)} \\ 
\times \mbox{P(wandering|We,had,been)}.$$

However, the likelihood that more and more words will occur next to each other in an identical sequence becomes smaller and smaller, making prediction difficult. Alternatively, we can approximate the probability of a word depending on only the previous word. This is known as the 'Markov assumption' and it allows us to approximate the probability using only the last $n$ words [@brown1992class]:

$$\mbox{Pr(We,had,been,wandering)} \approx \mbox{Pr(We)} \times \mbox{P(had|We)} \times \mbox{P(been|had)} \\ 
\times \mbox{P(wandering|been)}.$$

As a bi-gram model only considers the immediately preceding word, under the Markov assumption, an $n$-gram model can be reduced to a bi-gram model with $n$ being any number:

$$\mbox{Pr}(w_{n} | w^{n-1}_{1}) \approx \mbox{P}(w_{n} | w^{n-1}_{n - N+1}))$$
where $N$ is **(Ke-Li add)**.

Language models underpinned by n-grams are widely applied in text prediction, spelling-correction, and machine translation **(Ke-Li add citation)**. In our application, we use a tri-gram based model to detect both real-word and non-word errors and correct them with the candidate word that has the highest probability. **(TODO: Add more about how we use n-gram once the functions in the app are in place.)**

However, n-gram models do not take the linguistic structure of language into account. For instance, @rosenfeld2000two [p. X]  **(Ke-Li add)** discusses language in this context, saying that '...it may as well be a sequence of arbitrary symbols, with no deep structure, intention or thought behind'. The prediction of the next word is based on only a few preceding words and broader context is not taken into account. Hence next-word prediction using n-gram based language models can be limited.

## Word Embedding Models
Turning words into numeric representations is an important step for machine to understand and process words. The most basic way to turn a word into a vector is one-hot encoding. However, with one-hot encoding, the number of dimensions depends on the size of the vocabulary in a given corpus, and can easily grow to extremely large number of dimensions. Word embedding models work by representing words as numerical dense vectors so the dimensions are reduced by representing the words by features **(Keli add citation)**. Representing words as dense vectors has a long history, but the idea of a dense embedding space for words was first explored in 2003 by @bengio2003neural, and became the prominent driver for word embedding models. In the work of @bengio2003neural, words are encoded into feature vectors that represents different aspects of the word, and each word is associated with a point in a vector space. The words that are similar will have similar representations in a vector space. The similarity of the words, however, does not depend on their spelling or forms, but by their co-occurrence in given contexts. For example, a "cat" is more similar to a "dog" instead of a "car" because we are more likely to talk about cats and dogs in a same conversation. Therefore, the co-occurrence of "cat" and "dog" has higher probability than the co-occurrence of "cat" and a "car". In other words, the distributed representation is learned by the usage of words. This permits words that are used in similar contexts to result in having similar vector representations and their semantic meanings are captured naturally.

### Word2Vec

Developed by Google, Word2Vec is a word embedding model that uses shallow neural network. While it is not a deep learning model, it turns words into a vector representations that deep neural networks can understand. It takes a corpus as its input and output a set of feature vectors that represent the words in the given corpus. @mikolov2013efficient proposed two architectures for learning distributed representations of words within the framework of Word2Vec — Continuous Bag-of-Words Model (CBOW) and Continuous Skip-gram Model. Both architectures contains three layers: input layer, projection layer and output layer. The input layer encodes words within a context window to one-hot vectors. Using a projection matrix, the input layer is  projected to a projection layer and the features are. Finally, the output layer is a softmax regression classifier that obtains the posterior distribution of words (@rong2014word2vec). The input and output vectors differ in the two architectures. CBOW uses several surrounding words, or the context, to predict a target word. On the other hand, continuous skip-gram takes a single word as input to predict its neighbouring words. For example, consider the sentence "I think therefore I am", a famous quote by Descartes. The focus word is we want to learn the embedding is set to "therefore", and the context window is set to two which will give us the two immediately preceding words and two immediate proceeding words of the focus word. The input for the CBOW is "I, think, I, am" and the output is "therefore". In contrast, the skip-gram model will take "therefore" as the input and out "I, think, I, am" as the prediction.

Word2Vec is trained on the Google News data set that contains 6 billion tokens and the vocabulary size is restricted to one million most common words (@mikolov2013efficient). The enormous size of data set catches accurate statistical pattern of language and results a comprehensive mapping of word relations to a vector place. The positions of the word vectors reflect their semantic relationships between each other. Therefore, the Word2Vec model performs especially well in word analogy tasks through answering questions such as "$king$ is to $man$ as $queen$ is to __?" @pennington2014glove. The vectorization of words allows the operation of linear algebra for the word vectors in order to get the answer:

>  vector("King") - vector("Man") + vector("Woman") = vector("Queen")

Other Word2Vec use cases:
1. Find the most similar words (**Keli add example**)
2. Calculate the similarities between a pair of words (**Keli add example**)
3. Find the odd word in a list of words (**Keli add example**)

### GloVe

GloVe stands for global vectors for word representation. It is a pre-trained word embedding model that can capture semantic meanings of words developed by Stanford. In the Word2Vec model, the model learn word presentations locally within a context window. On the other hand, GloVe considers the properties of the corpus globally. The benefit of the global approach is that the model can learn about the global statistics of an entire corpus. GloVe can then obtain the co-occurrence probabilities by creating a co-occurrence count matrix for the entire training data set. However, GloVe goes beyond calculating distribution of word by the probability of co-occurrence; it reveals properties of words by the ratio of the probability of co-occurrence of words. For example, consider the two words we want to obtain the word vectors from "ice" and "steam". GloVe reveals their properties by examining their co-occurrence with a variety of probe words, $k$. $K$ can be "solid", which is the property of ice; "gas", which is the property of steam; and "water", which is related to both "ice" and "steam". The ratio of co-occurrence probability is calculated as the following:

> P(k|ice)/P(k|steam)

The ratio between the co-occurrence probability of "ice, solid" pair and "steam, solid" pair is high because the co-occurrence of "ice" and "solid" is higher than the co-occurrence of "steam" and "solid". Reversely, the ratio between the co-occurrence of "ice, gas" pair and "steam, gas" pair is low. On the other hand, the ratio of co-occurrence probability of "ice, water" pair and "steam, water" pair are close to 1 because water is relevant to both "ice" and "steam". Probability alone does not distinguish well relevant words from irrelevant words. For this reason, the researcher of GloVe argued that "the appropriate starting point for word vector learning should be with ratios of co-occurrence probabilities rather than the probabilities themselves (@pennington2014glove)."

## Pre-trained transformer language models

In recent years, the use of pre-trained language models has played a significant role in advancing natural language processing (NLP) tasks such as reading comprehension, text generation, and even creative writing **(Ke-Li add ref)**. Pre-trained models remove the need for researchers to prepare training datasets or allot computational resources to the stage of pre-training. Because of this, there has been substantial development in NLP research. OpenAI and Google are two companies that are considered pioneers in artificial intelligence and they both develop pre-trained models that are considered state-of-art; we will discuss the GPT models from OpenAI and the BERT model from Google.

### GPT

Draw on and cite: 'Language Models are Few-Shot Learners'

OpenAI is an AI research company who openly pursues Artificial General Intelligence—to have machines learn and understand any intellectual tasks as human do **(Ke-Li add ref)**. OpenAI's Generative Pre-Trained Transformer (GPT) models are pre-trained language models that are built based on the transformer architecture. Proposed by @vaswani2017attention from Google in 2017, the Transformer is a novel network architecture for neural network that outperforms RNN-based models in computational efficiency and CNN-based in (something that it doesn't have to look for longer sentences) **(Keli add citation)**. [Write about self-attention layer and how it differentiate from RNN and CNN]. [Paraphrase: The transformer is based solely on attention mechanisms, dispensing with recurrence and convolutions entirely, and has proven extremely effective compared to its predecessor the LSTM].

As of August 2020, OpenAI has three generations of GPT models. GPT was introduced to the public in June, 2018. In the year 2019, its successor GPT-2 was released in three stages. Not long after, in June 2020, OpenAI announced GPT-3, which is the latest GPT model as of July 2020 (@brown2020language). Since the API of GPT-3 at the moment is only available to a few invited users, public applications of GPT-3 are limited. On the other hand, its predecessor, GPT-2, has enabled researchers made breakthroughs in the past years. @alt2019fine extended GPT-2 to relation extraction, which is a task of identifying the relationship between concepts appeared in text. They had shown a milestone in predicting larger range of distinct relation types with higher confidence **(Ke-Li add ref)**. In the field of speech recognition, data scarcity is often an issue due to the difficulty to collect large amount of data from human for learning **(Ke-Li add ref)**. @bird2020lstm employed GPT-2 as part of the solution to generate augmented data for classification. GPT-2 also helped NLP researchers tackling the task of textual paraphrasing **(Ke-Li add ref)**. @hegde2020unsupervised used GPT-2 to build an unsupervised paraphrasing model that is not restricted by domain shift within an article, which is a problem faced by the usual supervised models of paraphrasing.

Most of prior language models are trained in a single domain by supervised learning to perform a single task. OpenAI, on the other hands, train the GPT models to perform many tasks (@radford2019language). GPT-2 is pre-trained on WebText, a comprehensive data set that is scrapped from millions of webpages and contains total of 40 GB of text (@radford2019language). WebText contains all kinds of textual data: dialogues, comments, translations, questions and answers...etc. Such diversity is critical for GPT-2 to be a generalized model that can perform multiple tasks with any domain boundary. However, GPT-2 has the capacity to adapt to a context and generates custom output with the "fine-tuning" technique **(Ke-Li add ref)**. While the pre-training stage is unsupervised and mandatory, fine-tuning is an optional training stage where GPT can learn to perform new tasks in a supervised or reinforced approach (@ziegler2019fine) **(Ke-Li add examples)**. During the pre-training stage, the model is trained to perform text completion task. Give it a prompt to start and it will generate texts that continues what has been started. In the fine-tuning stage, @radford2019language provided additional training data sets for GPT-2 to generate the texts with specific sentiment, or be descriptive, or even summarize the prompt text.

In our project, we aim to use fine-tune GPT-2 to perform a text cleaning task **(Ke-Li add more about how we use fine-tuning to train GPT-2 for text cleaning)**.

A language model is just a probability:
> p(output | input)

To model all NLP tasks, we cab represent them with one extra conditioning:
> p(output | input, task)

OpenAI has made GPT-2 open-sourced through GitHub with Python and TensorFlow. There is also a package in R that wraps the Python code into a R package, created by **(Keli add citation)**. However, the R package only permits exploratory experiments with the pre-trained tasks; fine-tuning for new tasks is not available thorough R.

GPT-3 has over 175 billions of parameters (@brown2020language). The parameters refer to the weights of the connections that enable the neural network to learn **(Keli add citation)**. In the past few years, the growth of parameters of transformer language models expands rapidly. The first generation GPT has 110 million parameters. Released in 2018, GPT-2 has 1.6 billion parameters, and within two years, GPT-3 has grown the size to 175 billion, which is over 100 times than its predecessor. The more parameters there are the better the model can generalize to new tasks; therefore, the increasing amount of parameters accelerates the improvements in text synthesis and downstream NLP tasks **(Keli add citation)**. GPT-3 can learn NLP tasks with just a handful of example; OpenAI refers this ability as few-shot learning, or in-context learning. **(Keli add citation)**. As a result, we only need to provide GPT-3 a few demonstrations for it to learn a new NLP task. For example, OpenAI team trained GPT-3 to generate "news articles". The prompt input is a plausible first sentence of a news story written by human, and the output is an entire news article generated by the model itself. Without the demonstrations, GPT-3 tends to treat the prompts as "tweets" and generate texts that serve as responses or follow-up tweets. The team fed a few previous news articles in the model’s context as demonstrations, and the model learned to produce the outputs that adapt to the style and length of news stories (@brown2020language). This few-shot learning ability indicates that the model can learn like human do and would democratize the use of language models to broader fields.

### BERT

BERT stands for Bidirectional Encoder Representations from Transformers (@devlin2018bert). It was released by researchers at Google AI Language in 2018. The key innovation of BERT is its application of bidirectional training of the Transformer model, which results in a deeper sense of language context and flow. What is bidirectional training? In the context of English language, a sentence is constructed from left to right. The context on the left is used to determine the next word on the right. In this sense, GPT models are unidirectional since the probability of the next word is dependent only on the past. It is also possible to reverse the direction and model the sentence from right to left, and BERT incorporates both directions of left-to-right and right-to-left to model bidirectional contexts. BERT uses the bidirectional approach to alleviate by "masked language model" (MLM). The MLM model masks certain words in the input corpus and predict the original words that are masked. The objective is to fuse the representation from the left and right context to enable the pre-training of the deep bidirectional Transformer. For example:

Because BERT employs the Transformer architecture with the self-attention mechanism, fine-tuning is also available.

# Stylized example

**(TODO: This stylized example need to be changed and updated to actually use GPT-3 at some point.)**

In our application, we deploy OpenAI GPT-3 to generate text and ...

As a stylized example, let's consider the following actual paragraph from Jane Eyre, by Charlotte Bronte:

> There was no possibility of taking a walk that day. We had been wandering, indeed, in the leafless shrubbery an hour in the morning; but since dinner (Mrs. Reed, when there was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, that further out-door exercise was now out of the question.

Let's pretend that this text had been created from optical character recognition and that it had the following errors: some 'h' were replaced with 'b'; and some 'd' have been replaced with 'al':

> There was no possibility of taking a walk that day. We had been wandering, indeed, in tbe leafless shrubbery an hour in tbe morning; but since dinner (Mrs. Reeal, when tbere was no company, dined early) the cold winter wind had brought with it clouds so sombre, and a rain so penetrating, tbat further out-door exercise was now out of the question.

Assume a model that is trained to perfectly forecast the next word in Jane Eyre. For this fragment there are 62 words, comprising 5 errors and 57 correct words. So the internal consistency score of this fragment would be: 57/62 = 0.919. When we recognise and correct the errors, this consistency score would increase to 1.

Similarly, assume a model that is trained on an external data source. This means that it will recognise the the cases where some 'h' were replaced with 'b', but not recognise that 'Reed' has become 'Reeal'. Hence, the external consistency score would be 58/62 = 0.935.

# Data

Various data can be used....

# Application

# Discussion

Internal validity vs external- one is their own words the other is a general set of words.

In the same way that precision and recall provide important measures…

# Appendix

## Literature summaries

(These are just here for now - we'll remove later.)

### Article: Improving Language Understanding by Generative Pre-Training (2018)
Introduction of a semi-supervised approach for language understanding machine learning tasks. The approach is the combinaiton of unsupervised pre-training and supervised fine-tuning.

### Article: Language Models are Few-Shot Learners (2020)
Official paper from OpenAI to introduce OpenAI GPT-3. Emphasizes on the "fewer-shots" and "task-agnostic" aspects of the latest langual model compared to its predecessor GPT-2.

## Discussion of how GPT-3 works

**(Ke-Li - it would be good for us to go through how it actually works in as much detail as we can manage (in our own words). This is mostly just to force us to learn how it works and may be taken out of the final version of the paper, but it's still an important exercise.)**

# References