### 2020-10-04 outputs/documents/notes/paper_code_documentation.Rmd
code trial for text infilling paper

---
title: "paper_code_documentation"
author: "Keli Chiu"
date: "03/10/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Paper: Text Infilling

Published date: January 18, 2019

Authors: Wanrong Zhu, Zhiting Hu, Eric P. Xing

Affiliation: Peking University, Carnegie Mellon University, Petuum Inc.

Code URL: https://github.com/VegB/Text_Infilling

### Result

Not able to follow through the code tutorial.

### Steps:

1. Download the code and install Texar.

``` {python, eval = FALSE}
git clone https://github.com/VegB/Text_Infilling
cd Text_Infilling
pip install --user -e .    
```

2. Navigate to the infilling demo directory and download the dataset. The instruction provided on Github didn't work, I followed the `README.MD` in the `text_infilling` directory from this point. Downloaded data is stored in `yelp_data/`.

``` {python, eval = FALSE}
cd text_infilling
python data_utils.py
```

You might have to run extra code to install required packages if you don't have them already. In my case, I had to get `tensorflow` and `matplotlib`.

``` {python, eval = FALSE}
pip install "tensorflow>=1.15,<2.0" 
pip install matplotlib
```

3. Code template to execute is provided as follow:

``` {python, eval = FALSE}
python [MODEL].py --mask_rate [MASK_RATE] --blank_num [BLANK_NUM] 
  --filename_prefix 'pos.' --data_dir './yelp_data/pos/'
```

Here is what I tried, set mask rate to 30%, number of blank in an example is 1, and chose the self-attention mode.

``` {python, eval = FALSE}
python self_attn.py --mask_rate 30 --blank_num 1 --filename_prefix 'pos.' 
  --data_dir './yelp_data/pos/'
```

An Error was raised from the above code. The error message:

``` {python, eval = FALSE}
Traceback (most recent call last):
  File "self_attn.py", line 335, in <module>
    tf.app.run(main=_main)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/tensorflow_core/python/platform/app.py", 
  line 40, in run
    _run(main=main, argv=argv, flags_parser=_parse_flags_tolerate_undef)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/absl/app.py", 
  line 300, in run
    _run_main(main, args)
  File "/Users/kelichiu/Desktop/project_tif/lib/python3.7/site-packages/absl/app.py", 
  line 251, in _run_main
    sys.exit(main(argv))
  File "self_attn.py", line 64, in _main
    hparams=args.word_embedding_hparams)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedders.py", 
  line 53, in __init__
    self._hparams)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedder_base.py", 
  line 33, in _init_parameterized_embedding
    hparams, init_value, num_embeds, self.variable_scope)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/modules/embedders/embedder_utils.py", 
  line 191, in get_embedding
    initializer = layers.get_initializer(hparams["initializer"])
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/core/layers.py", 
  line 333, in get_initializer
    initializer = utils.get_instance(hparams["type"], kwargs, modules)
  File "/Users/kelichiu/Desktop/repos/Text_Infilling/texar/utils/utils.py", 
  line 209, in get_instance
    (class_.__module__, class_.__name__, key, class_args))
ValueError: Invalid argument for class 
  tensorflow.python.ops.init_ops.RandomNormal: mean, valid args:set()
```

I have change the --mask_rate 30 and --blank_num 1 with variations between int, float and percentage, but no value could successfully run the code. Googled solutions but no relevant answers were found.

## Paper: Text Infilling

Extra step
pip install nltk